Simple connectome inference from partial correlation statistics in calcium imaging

### 2014-06-28 article.tex
AQ : minor changes in appendix

In Section~\ref{sec:filter}, we introduce four filtering functions ($f$, $g$,
$h$, and $w$) that are composed in sequence (i.e., $w \circ h \circ g \circ
f$) to provide the signals from which to compute partial correlation
statistics.

Filtering is modified as follows in the tuned method:
\begin{itemize}
\item In addition to $f_1$ and $f_2$ (Equations \ref{eq:symetric-median} and
  \ref{eq:weighted-asymetric-median}), two alternative low-pass filters $f_3$
  and $f_4$ are considered:
\begin{align}
f_3(x^t_i) &= x^{t-1}_i + x^{t}_i + x^{t+1}_i + x^{t+2}_i, \label{eq:asymetric-median-forward} \\
f_4(x^t_i) &=  x_i^t + x^{t+1}_i  + x^{t+2}_i + x^{t+3}_i. \label{eq:asymetric-median}.
\end{align}
\item An additional filter $p$ is applied to smoothen differences in peak magnitudes
  that might remain after the application of the hard-threshold filter $h$:
\begin{align}
p(x^t_i) = (x_i^t)^c,
\end{align}
with $c=0.9$.
\item Filter $w$ is replaced by a more complex filter $w^*$ defined as:
\begin{align}
 w^*(x^{t}_i) &= {(x^{t}_i + 1 )^{\left (1 + \frac{1}{\sum_{j} x^{t}_j}\right )}}^{k(\sum_{j} x^{t}_j)}
\end{align}
where the function $k$ is a piecewise linear function optimized separately for
each filter $f_1$, $f_2$, $f_3$ and $f_4$ (see the implementation for full
details). Filter $w$ in the simplified method is a special case of $w^*$ with
$k(\sum_j x_j^t)=1$.
\end{itemize}
The pre-processed time-series are then obtained by the application of the
following function: $w^*\circ p \circ h \circ g \circ f_i$ (with $i=1$,2,3, or 4).

@@

Two issues remain with this heuristic. First, it does not allow to detect a
directed relationship if both neurons are activated simultaneously at the same
time step, ie., when the delay between the two activations is below the time
resolution. Second, one needs to compare $s_{j,i}$ and $s_{i,j}$ to decide on
edge orientation. Indeed, edge orientation can only be predicted with some
confidence if either $s_{i,j}$ is greater than $s_{j,i}$ ($i\rightarrow j$) or
$s_{i,j}$ is lower than $s_{j,i}$ ($j\rightarrow i$). To address these issues,
we computed a differential pairwise coefficient $z_{ij}$ defined as follows:
\begin{align}
z_{ij} = \sum_{t=1}^{T - 1}
    \mathbb{1}((x_j^{t+1} - x_i^t) \in \left[\phi_1, \phi_2 \right]) -
    \mathbb{1}((x_i^{t+1} - x_j^t) \in \left[\phi_1, \phi_2 \right]),
\end{align}
where $T$ is the number of time steps, $\phi_1$ and $\phi_2$ are parameters.

\textcolor{red}{PG: je ne comprends pas ce qui est fait exactement. Mettre les
  maths? Valeurs de $\phi_1$ et $\phi_2$ ?} At the light of the poor results of this heuristic taken on its own
(on Table \ref{tab:directivity} in Section~\ref{sapp:results}) as
\textit{directivity}, we use it to make asymmetric the degree of association
matrix using partial correlation statistics. We expect that when the heuristic
is highly confident (i.e., $z_{i,j} \gg 0 \gg z_{j,i}$) the degree of
association of edge $(i,j)$ is increased while the degree of association
corresponding to the edge $(j,i)$ is reduced. Even if this adjustment is
subtle, it might have a slight positive effect on the ROC AUC.

>>>>

In Section~\ref{sec:filter}, we introduced four filtering functions ($f$, $g$,
$h$, and $w$) that are composed in sequence (i.e., $w \circ h \circ g \circ
f$) to provide the signals from which to compute partial correlation
statistics.

Filtering is modified as follows in the tuned method:
\begin{itemize}
\item In addition to $f_1$ and $f_2$ (Equations \ref{eq:symetric-median} and
  \ref{eq:weighted-asymetric-median}), two alternative low-pass filters $f_3$
  and $f_4$ are considered:
\begin{align}
f_3(x^t_i) &= x^{t-1}_i + x^{t}_i + x^{t+1}_i + x^{t+2}_i, \label{eq:asymetric-median-forward} \\
f_4(x^t_i) &=  x_i^t + x^{t+1}_i  + x^{t+2}_i + x^{t+3}_i. \label{eq:asymetric-median}.
\end{align}
\item An additional filter $p$ is applied to smoothen differences in peak magnitudes
  that might remain after the application of the hard-threshold filter $h$:
\begin{align}
r(x^t_i) = (x_i^t)^c,
\end{align}
with $c=0.9$.
\item Filter $w$ is replaced by a more complex filter $w^*$ defined as:
\begin{align}
 w^*(x^{t}_i) &= {(x^{t}_i + 1 )^{\left (1 + \frac{1}{\sum_{j} x^{t}_j}\right )}}^{k(\sum_{j} x^{t}_j)}
\end{align}
where the function $k$ is a piecewise linear function optimized separately for
each filter $f_1$, $f_2$, $f_3$ and $f_4$ (see the implementation for full
details). Filter $w$ in the simplified method is a special case of $w^*$ with
$k(\sum_j x_j^t)=1$.
\end{itemize}
The pre-processed time-series are then obtained by the application of the
following function: $w^*\circ r \circ h \circ g \circ f_i$ (with $i=1$,2,3, or 4).

@@

Two issues remain with this heuristic. First, it does not allow to detect a
directed relationship if both neurons are activated simultaneously at the same
time step, ie., when the delay between the two activations is below the time
resolution. Second, one needs to compare $s_{j,i}$ and $s_{i,j}$ to decide on
edge orientation. Indeed, edge orientation can only be predicted with some
confidence if either $s_{i,j}$ is greater than $s_{j,i}$ ($i\rightarrow j$) or
$s_{i,j}$ is lower than $s_{j,i}$ ($j\rightarrow i$). To address these issues,
we computed a differential pairwise coefficient $z_{i,j}$ defined as follows:
\begin{align}
z_{i,j} = \sum_{t=1}^{T - 1}
    \mathbb{1}((x_j^{t+1} - x_i^t) \in \left[\phi_1, \phi_2 \right]) -
    \mathbb{1}((x_i^{t+1} - x_j^t) \in \left[\phi_1, \phi_2 \right]),
\end{align}
where $T$ is the number of time steps, $\phi_1$ and $\phi_2$ are parameters.

\textcolor{red}{PG: je ne comprends pas ce qui est fait exactement. Mettre les
  maths? Valeurs de $\phi_1$ et $\phi_2$ ?} At the light of the poor results of this heuristic taken on its own as
\textit{directivity} 
(on Table \ref{tab:directivity} in Section~\ref{sapp:results}), we use it to asymmetrize the degree of association
matrix using partial correlation statistics. We expect that when the heuristic
is highly confident (i.e., $z_{i,j} \gg 0 \gg z_{j,i}$) the degree of
association of edge $(i,j)$ is increased while the degree of association
corresponding to the edge $(j,i)$ is reduced. Even if this adjustment is
subtle, it might have a slight positive effect on the ROC AUC.
